% IMPORTANT: You MUST compile this in Overleaf using the "XeLaTeX" engine. 
% (Menu -> Compiler -> XeLaTeX)

\documentclass[11pt, a4paper]{article}

\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{caption}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{float}
\usepackage{setspace}
\usepackage{makecell}
\usepackage{listings}
\usepackage{array}

% Language setup
\setmainlanguage{english}
\setotherlanguage{arabic}

% Font setup (Amiri is available on Overleaf)
\newfontfamily\arabicfont[Script=Arabic]{Amiri}

% Set page margins
\geometry{
    a4paper,
    top=2.5cm,
    bottom=2.5cm,
    left=2.5cm,
    right=2.5cm
}

% Hyperlink configuration
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=teal,
    citecolor=blue,
    pdftitle={Jawhar - Arabic Morphosyntactic Explorer Project Report},
}

% Colors
\definecolor{primaryGold}{HTML}{FFD33D}

\title{
    \vspace{2cm}
    \textbf{\huge Jawhar (\textarabic{Ш¬ЩҲЩҮШұ})} \\
    \vspace{0.5cm}
    \Large Arabic Morphosyntactic Explorer \\
    \vspace{1cm}
    \large Capstone Project Report
    \vspace{2cm}
}
\author{\textbf{Meher Ali} \\ Software Engineering Capstone Project}
\date{\vspace{1cm} \today}

\begin{document}

\maketitle
\thispagestyle{empty}
\newpage

\pagenumbering{roman}
\begin{abstract}
\noindent 
The Arabic language, particularly the Classical Arabic found in the Quranic corpus, possesses a highly complex morphosyntactic structure. Traditional tools for linguistic analysis are often disconnected, lacking either depth in Natural Language Processing (NLP) or an intuitive user interface for learners. This report details the development of \textbf{Jawhar (\textarabic{Ш¬ЩҲЩҮШұ})}, an AI-driven Arabic Morphosyntactic Explorer. Jawhar bridges the gap between advanced computational linguistics and pedagogical application by integrating CAMeL Tools for precise morphological parsing, ChromaDB for semantic retrieval operations, and Google's Gemini Flash AI architecture for context-aware tutoring. The system employs a cross-platform React Native frontend and a FastAPI backend deployed on Hugging Face Spaces. The resulting platform enables users to semantically search the corpus, interact deeply with word-level morphology (Sarf), and receive grammatical and pedagogical insights (I'rab) through a seamless, robust interface.
\end{abstract}

\tableofcontents
\newpage
\pagenumbering{arabic}
\setstretch{1.2}

\section{Introduction}

\subsection{Motivation}
Classical Arabic is a highly inflected, derivational language, where a vast vocabulary is generated from trilateral or quadrilateral root consonants through defined morphological templates (\textarabic{Awzan/ШЈЩҲШІШ§ЩҶ}). For students and researchers, decoding this morphology (Sarf) and the syntactic rules determining word endings (\textarabic{I'rab/ШҘШ№ШұШ§ШЁ}) is traditionally a steep learning curve. While tools exist that serve NLP researchers, few offer end-user interfaces dedicated to pedagogical improvement, semantic search capabilities, and instant AI tutoring.

\subsection{Project Objectives}
The Jawhar project was conceptualized with the following core objectives:
\begin{itemize}
    \item \textbf{Deep Linguistic Integration}: To utilize state-of-the-art NLP models (CAMeL Tools) to automatically extract root, lemma, part-of-speech (POS), voice, aspect, and gender from raw Arabic texts.
    \item \textbf{Semantic Search (RAG)}: To supersede traditional keyword search by embedding the semantic meaning of verses using a vector database (ChromaDB), allowing users to find verses by conceptual queries.
    \item \textbf{Pedagogical AI Tutor}: To implement a custom Gemini-powered AI agent wrapped in an intuitive chat interface capable of explaining morphological outputs in natural language.
    \item \textbf{Cross-platform Accessibility}: To build a single unified application (React Native / Expo) that supports native Android deployments as well as web, maintaining precise Right-To-Left (RTL) capabilities.
\end{itemize}

\section{System Architecture}

Jawhar leverages a modular, separated frontend-to-backend architecture, optimizing for computational efficiency on the NLP front and responsive interactions on the mobile front.

\subsection{Architectural Components}
The system architecture encompasses three primary layers:

\begin{enumerate}
    \item \textbf{Frontend Application (Presentation Layer)}: Developed using React Native with the Expo framework. It natively provisions bidirectional text flow and an adaptive UI.
    \item \textbf{FastAPI Backend (Service Layer)}: A Python 3.12+ REST API application serving as a highly concurrent gateway orchestration layer.
    \item \textbf{NLP \& AI Microservices (Processing Layer)}:
    \begin{itemize}
        \item \textit{CAMeL Tools}: Handles the heavy-lifting of root extraction, diacritization matching, and detailed metadata generation.
        \item \textit{Gemini Flash}: Serves dual purposes---generating pedagogical explanations of linguistic structures and creating vector embeddings for the text.
        \item \textit{ChromaDB}: Persistent local vector store optimizing k-Nearest Neighbors (k-NN) cosine similarity search for the Quranic dataset.
    \end{itemize}
\end{enumerate}

\section{Dataset Description}

\subsection{Quranic Corpus}
The system operates on a comprehensive Quranic dataset containing the complete text of the Quran (6,236 verses across 114 chapters). Each verse entry includes structured fields as presented in Table~\ref{tab:dataset-fields}.

\begin{table}[H]
\centering
\caption{Quranic Dataset Fields}
\label{tab:dataset-fields}
\begin{tabular}{|l|l|p{5cm}|}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
surah\_no & Integer & Chapter number (1вҖ“114) \\
surah\_name\_en & String & English chapter name \\
surah\_name\_ar & String & Arabic chapter name \\
surah\_name\_roman & String & Transliterated name \\
ayah\_no\_surah & Integer & Verse number within chapter \\
ayah\_no\_quran & Integer & Global verse number \\
ayah\_ar & String & Arabic verse text \\
ayah\_en & String & English translation \\
list\_of\_words & Array & Tokenized word array \\
place\_of\_revelation & String & Meccan or Medinan \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data Characteristics}
The dataset exhibits several characteristics relevant to NLP processing:
\begin{itemize}
    \item \textbf{Bilingual Content}: Each verse provides both Arabic source text and English translation, enabling bilingual analysis.
    \item \textbf{Tokenization}: Words are pre-tokenized, facilitating word-level analysis.
    \item \textbf{Metadata Enrichment}: Verse origin (Meccan/Medinan), chapter position, and verse statistics are included.
\end{itemize}

\section{Technical Implementation}

\subsection{Morphological Analysis Engine}
Sarf (Morphology) is the foundational pivot of Jawhar. The \texttt{MorphologyService} in the backend isolates the \texttt{camel\_tools} dependency. When a user queries a single word or an entire verse, the toolkit evaluates probabilities and returns the highest likelihood token breakdown. 
A key implementation success was the normalization logic applied on the frontend, where diacritics and esoteric symbols (e.g., Quranic stop marks) are stripped to match touch events on the screen accurately to the parsed NLP tokens returning from the robust backend structure.

\subsection{Semantic Retrieval-Augmented Generation (RAG)}
Semantic search is powered by \texttt{ChromaDB}. Upon bootstrapping, the \texttt{VectorStore} embeds the initial Quranic chapters (via Google GenAI embedding models). Queries initiated by users are vectorized, and the top-$K$ verses possessing the highest semantic similarity are returned, enabling interactions like searching for ``Verses about patience in hardship.''

\subsection{AI Pedagogical Tutor}
Jawhar integrates the \texttt{gemini-3.1-flash-lite-preview} model. In the \textbf{Verse Detail Screen}, selecting a word performs an instant, dynamic fetch over REST API requesting a targeted grammatical explanation. The API engineers a highly structured prompt forcing the LLM to analyze the CAMeL output from a pedagogical lens, strictly formatting responses into Sarf (morphology), I'rab (syntax), and Summary segments.

\subsection{Frontend Engineering \& RTL Compliance}
React Native Expo Router manages deep linking. A critical UI challenge was preserving standard application patterns (like back functionality) while forcing Right-To-Left directionality. Utilizing Expo's \texttt{I18nManager.forceRTL(true)}, layouts adapt flawlessly, keeping Arabic primary while aligning corresponding English translations logically. Custom Glassmorphism styles and specific ``Amiri'' typography are loaded statically to convey an elegant, modern visual identity.

\subsection{API Endpoints}
The backend exposes RESTful endpoints for all core functionality. Table~\ref{tab:api-endpoints} summarizes the available endpoints.

\begin{table}[H]
\caption{REST API Endpoints}
\label{tab:api-endpoints}
\begin{tabular}{|l|l|p{6cm}|}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Description} \\
\midrule
GET & /api/v1/verses/\{chapter\}/\{verse\} & Retrieve specific verse \\
GET & /api/v1/verses/\{chapter\} & List chapter verses with pagination \\
POST & /api/v1/verses/search & Semantic verse search \\
GET & /api/v1/morphology/\{ch\}/\{verse\} & Get verse morphological analysis \\
POST & /api/v1/morphology/analyze & Analyze user input text \\
POST & /api/v1/morphology/analyze-word & Single word with AI explanation \\
POST & /api/v1/chat/ & AI tutoring conversation \\
\bottomrule
\end{tabular}
\end{table}

\section{Deployment \& CI/CD Operations}

To guarantee stable and consistent deployment, the project adopted a hybrid release paradigm:

\begin{itemize}
    \item \textbf{Backend Hosting}: The FastAPI environment, which exceeds standard limits due to dense NLP library models, is containerized via a custom Dockerfile and deployed via GitHub workflows to \textbf{Hugging Face Spaces}. A separate deployment branch (\texttt{hf-deploy}) maintains Hugging Face specific YAML metadata configurations for automated Docker builds.
    \item \textbf{Frontend Application}: Expo Application Services (EAS) were configured via \texttt{eas.json} to decouple local builds. The final deliverable is compiled as a standalone \texttt{.apk} (Android Package) to be distributed through the GitHub Releases tab, allowing direct end-user installation without local SDK friction.
\end{itemize}

\section{Results and Performance}

The development successfully united complex computational linguistics with mobile engineering. 
\textbf{Key accomplishments include}:
\begin{itemize}
    \item Centralized morphological normalization: Users observe over 12 detailed attributes per word (Voice, Aspect, Mood, Gender, Root, Lemma, etc.) interchangeably seamlessly between search, deep verse analysis, and free text parsing.
    \item Deployment stability and 0-configuration orchestration: The provided \texttt{dev.sh} spins up virtual environments across Python and Node.js consecutively in under 45 seconds for parallel local development.
\end{itemize}

\subsection{Functional Implementation Status}
Table~\ref{tab:feature-status} presents the implemented features and their operational status.

\begin{table}[H]
\caption{Implemented Features and Status}
\label{tab:feature-status}
\begin{tabular}{|lcl|}
\toprule
\textbf{Feature} & & \textbf{Status} \\
\midrule
Morphological Analysis (Sarf) & $\checkmark$ & Operational via CAMeL Tools \\
Semantic Search (RAG) & $\checkmark$ & Operational via ChromaDB \\
AI Pedagogical Tutor & $\checkmark$ & Operational via Gemini 3.1 Flash \\
Bilingual UI (Arabic/English) & $\checkmark$ & Operational with RTL support \\
Mobile Application (Android) & $\checkmark$ & Built via Expo EAS \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Key Features}
The system implements five core features:

\begin{enumerate}
    \item \textbf{Deep Morphological Analysis (Sarf)}: Word-by-word breakdown showing root, lemma, POS tags, voice, aspect, gender, number, person, and grammatical case.
    \item \textbf{Semantic Search}: Vector-based similarity search enabling natural language queries such as "Verses about mercy and forgiveness."
    \item \textbf{AI Pedagogical Tutor}: Conversational interface providing explanations grounded in Quranic text with bilingual support.
    \item \textbf{Bilingual RTL Interface}: Complete Arabic/English support with automatic RTL/LTR layout switching.
    \item \textbf{Modern UI}: Dark mode theme with Glassmorphism design elements.
\end{enumerate}

\section{Limitations \& Future Work}

While foundational architecture is robust, the computational size of CAMeL databases dictates high memory constraints, pushing latency up slightly during cold starts on free-tier cloud containers. 
Future roadmap efforts should focus on:
\begin{enumerate}
    \item A fully offline mode for embedding models utilizing smaller ONNX-compiled equivalents (e.g., MiniLM) directly on Android.
    \item Expanding the vector database ingestion to cover the entire corpus rather than isolated selections limit due to rapid prototyping.
    \item Migrating the AI generation parameters to continuous streaming protocols to improve perceived loading thresholds for I'rab explanations.
\end{enumerate}

\section{Conclusion}

Jawhar stands as a successful Capstone project proving modern LLM and heuristic NLP methodologies can run conjointly inside scalable, user-centric mobile applications. By mitigating the computational convolution of Arabic linguistic tools into clean graphical interfaces and natural language chats, it establishes an exceptional pedagogical pipeline designed to modernize traditional Semitic linguistic studies.

\section{References}

\begin{enumerate}
    \item \textbf{CAMeL Tools}. (2024). CAMeL Tools: Arabic Natural Language Processing Tools. Carnegie Mellon University Qatar. Retrieved from \url{https://camel-tools.com/}
    
    \item \textbf{Google}. (2024). Gemini API Documentation. Google AI Studio. Retrieved from \url{https://ai.google.dev/}
    
    \item \textbf{ChromaDB}. (2024). Chroma: The AI-native open-source vector database. Retrieved from \url{https://www.trychroma.com/}
    
    \item \textbf{Tiangio }. (2024). FastAPI. Retrieved from \url{https://fastapi.tiangolo.com/}
    
    \item \textbf{Meta}. (2024). React Native. Retrieved from \url{https://reactnative.dev/}
    
    \item \textbf{Expo}. (2024). Expo: The fastest way to build React Native apps. Retrieved from \url{https://expo.dev/}
    
    \item \textbf{Hugging Face}. (2024). Hugging Face Spaces: ML demo hosting. Retrieved from \url{https://huggingface.co/spaces}
    
    \item \textbf{SRQ21 Collection}. (2024). The Siraj Quran 21 Corpus. Retrieved from \url{https://github.com/}
    
    \item \textbf{Python Software Foundation}. (2024). Python Programming Language (Version 3.12). Retrieved from \url{https://www.python.org/}
    
    \item \textbf{Node.js Foundation}. (2024). Node.js JavaScript Runtime. Retrieved from \url{https://nodejs.org/}
\end{enumerate}

\appendix
\section{System Architecture Diagram}

Figure~\ref{fig:system-architecture} presents the high-level system architecture.

\begin{figure}[H]
\centering
\caption{System Architecture Overview}
\label{fig:system-architecture}
\begin{scriptsize}
\begin{verbatim}
в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ
в”Ӯ                  Frontend (React Native/Expo)                в”Ӯ
в”Ӯ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ  в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ в”Ӯ
в”Ӯ  в”ӮAnalyzerв”Ӯ  в”ӮExplore в”Ӯ  в”Ӯ Chat    в”Ӯ  в”ӮSettings в”Ӯ в”Ӯ
в”Ӯ  в”Ӯ Tab   в”Ӯ  в”Ӯ  Tab   в”Ӯ  в”Ӯ  Tab   в”Ӯ  в”Ӯ  Tab    в”Ӯ в”Ӯ
в”Ӯ  в””в”Җв”Җв”Җв”Җв”¬в”Җв”Җв”ҳ  в””в”Җв”Җв”Җв”Җв”¬в”Җв”Җв”ҳ  в””в”Җв”Җв”Җв”Җв”¬в”Җв”Җв”ҳ  в””в”Җв”Җв”Җв”Җв”¬в”Җв”Җв”Җв”ҳ в”Ӯ
в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ
        в”Ӯ           в”Ӯ           в”Ӯ          в”Ӯ
        в””в”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҙв”Җв”Җв”Җв”Җв”Җв”¬в”Җв”Җв”Җв”Җв”Җв”ҙв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”ҳ
                       в”Ӯ
              в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв–јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ
              в”Ӯ  FastAPI    в”Ӯ
              в”Ӯ  REST API   в”Ӯ
              в””в”Җв”Җв”Җв”Җв”Җв”Җв”¬в”Җв”Җв”Җв”Җв”Җв”ҳ
                     в”Ӯ
    в”Ңв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”јв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”Җв”җ
    в”Ӯ                в”Ӯ               в”Ӯ
в”Ңв”Җв”Җв–јв”Җв”Җв”Җв”җ      в”Ңв”Җв”Җв”Җв”Җв–јв”Җв”Җв”Җв”җ    в”Ңв”Җв”Җв–јв”Җв”Җв”Җв”Җв”җ
в”ӮMorph в”Ӯ      в”ӮGemini  в”Ӯ    в”Ӯ RAG   в”Ӯ
в”ӮServiceв”Ӯ      в”ӮServiceв”Ӯ    в”ӮPipeline
в”Ӯ      в”Ӯ      в”Ӯ       в”Ӯ    в”Ӯ       в”Ӯ
в””в”Җв”Җв”¬в”Җв”Җв”Җв”ҳ      в””в”Җв”Җв”Җв”Җв”¬в”Җв”Җв”Җв”ҳ    в””в”Җв”Җв”¬в”Җв”Җв”Җв”Җв”ҳ
   в”Ӯ               в”Ӯ          в”Ӯ
в”Ңв”Җв”Җв–јв”Җв”Җв”җ      в”Ңв”Җв”Җв–јв”Җв”Җв”Җв”җ  в”Ңв”Җв”Җв–јв”Җв”Җв”Җв”җ
в”ӮCAMeL в”Ӯ      в”ӮGenAI в”Ӯ  в”ӮChromaв”Ӯ
в”ӮTools в”Ӯ      в”ӮEmbed в”Ӯ  в”ӮDB    в”Ӯ
в””в”Җв”Җв”Җв”Җв”Җв”Җв”ҳ      в””в”Җв”Җв”Җв”Җв”Җв”Җв”ҳ  в””в”Җв”Җв”Җв”Җв”Җв”Җв”ҳ
\end{verbatim}
\end{scriptsize}
\end{figure}

\section{Directory Structure}

Figure~\ref{fig:directory-structure} presents the project directory structure.

\begin{figure}[H]
\centering
\caption{Project Directory Structure}
\label{fig:directory-structure}
\begin{scriptsize}
\begin{verbatim}
jawhar/
в”ңв”Җв”Җ backend/                 # FastAPI application
в”Ӯ   в”ңв”Җв”Җ app/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ api/          # API routes
в”Ӯ   в”Ӯ   в”Ӯ   в””в”Җв”Җ routes/
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ verses.py
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ morphology.py
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ chat.py
в”Ӯ   в”Ӯ   в”Ӯ       в””в”Җв”Җ dictionary.py
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ core/         # Configuration
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ models/       # Pydantic models
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ services/     # Business logic
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ morphology.py
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ gemini.py
в”Ӯ   в”Ӯ   в”Ӯ       в”ңв”Җв”Җ rag.py
в”Ӯ   в”Ӯ   в”Ӯ       в””в”Җв”Җ vector_store.py
в”Ӯ   в”Ӯ   в””пҝҪпҝҪпҝҪв”Җ utils/       # Helpers
в”Ӯ   в”Ӯ   в””в”Җв”Җ main.py
в”Ӯ   в”ңв”Җв”Җ tests/           # Pytest tests
в”Ӯ   в”ңв”Җв”Җ scripts/        # Database population
в”Ӯ   в”ңв”Җв”Җ requirements.txt
в”Ӯ   в””в”Җв”Җ Dockerfile
в”ңв”Җв”Җ frontend/            # Expo React Native
в”Ӯ   в”ңв”Җв”Җ app/          # Expo Router screens
в”Ӯ   в”Ӯ   в””в”Җв”Җ (tabs)/
в”Ӯ   в”Ӯ       в”ңв”Җв”Җ index.tsx
в”Ӯ   в”Ӯ       в”ңв”Җв”Җ analyzer.tsx
в”Ӯ   в”Ӯ       в”ңв”Җв”Җ explore.tsx
в”Ӯ   в”Ӯ       в”ңв”Җв”Җ chat.tsx
в”Ӯ   в”Ӯ       в””в”Җв”Җ settings.tsx
в”Ӯ   в”ңв”Җв”Җ src/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ components/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ services/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ hooks/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ utils/
в”Ӯ   в”Ӯ   в”ңв”Җв”Җ types/
в”Ӯ   в”Ӯ   в””в”Җв”Җ constants/
в”Ӯ   в”ңв”Җв”Җ package.json
в”Ӯ   в””в”Җв”Җ app.json
в”ңв”Җв”Җ dataset/           # Quranic corpus
в”Ӯ   в””в”Җв”Җ quran_dataset.csv
в”ңв”Җв”Җ docs/             # Documentation
в”Ӯ   в””в”Җв”Җ jawhar_project_report.tex
в”ңв”Җв”Җ chroma_data/      # Vector database
в”ңв”Җв”Җ AGENTS.md
в””в”Җв”Җ README.md
\end{verbatim}
\end{scriptsize}
\end{figure}

\section{Technology Stack Summary}

Table~\ref{tab:tech-stack} provides a summary of the technology stack used in the project.

\begin{table}[H]
\caption{Technology Stack Summary}
\label{tab:tech-stack}
\begin{tabular}{|l|l|p{5cm}|}
\toprule
\textbf{Category} & \textbf{Technology} & \textbf{Purpose} \\
\midrule
Backend Framework & FastAPI & REST API service \\
NLP Library & CAMeL Tools & Arabic morphological analysis \\
Vector Database & ChromaDB & Semantic search storage \\
AI Model & Gemini Flash & Generative AI tutoring \\
Frontend & React Native (Expo) & Mobile application \\
Deployment & Hugging Face Spaces & Backend hosting \\
Language (BE) & Python 3.12 & Backend implementation \\
Language (FE) & TypeScript & Frontend implementation \\
\bottomrule
\end{tabular}
\end{table}

\section{Glossary}

\begin{table}[H]
\caption{Glossary of Terms}
\label{tab:glossary}
\begin{tabular}{|l|p{10cm}|}
\toprule
\textbf{Term} & \textbf{Definition} \\
\midrule
Sarf (Ш§Щ„ШөШұЩҒ) & Arabic morphological science dealing with word formation and root derivation \\
\midrule
I'rab (Ш§Щ„ШҘШ№ШұШ§ШЁ) & Arabic grammatical case marking indicating syntactic function \\
\midrule
Wazan (ЩҲШІЩҶ) & Morphological pattern applied to roots to generate words \\
\midrule
Root (Ш¬Ш°Шұ) & The consonantal base of an Arabic word (typically 3 or 4 letters) \\
\midrule
POS & Part-of-speech classification (Noun, Verb, Particle) \\
\midrule
RAG & Retrieval-Augmented Generation - AI architecture combining search with generation \\
\midrule
RTL & Right-To-Left text direction (Arabic script) \\
\midrule
LTR & Left-To-Right text direction (English script) \\
\midrule
ChromaDB & Vector database for semantic similarity search \\
\midrule
CAMeL Tools & Arabic NLP library from Carnegie Mellon Qatar \\
\midrule
Expo & Framework for building React Native apps \\
\midrule
APK & Android Package - compiled Android application \\
\midrule
FastAPI & Python web framework for REST APIs \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
